---
title: "Lab 7: SQL Murder - Data types"
author: "Karla Buitrago-Yepes"
format: html
editor: source
markdown:
execute: 
  echo: true
---

For this lab, you will be joining and filtering related data sets to
solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that
says `"Your name here!"`. Change this to be your name and save your file
either by clicking on the blue floppy disk or with a shortcut (command /
control + s).

Be sure to [commit the files to your
repo](https://happygitwithr.com/existing-github-first#stage-and-commit).

Let's get started!

# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid
    "object junk" in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users'
database skills, so they created a murder mystery. Can you solve this
crime in SQL City??

The relational data you will be working with contains tables with
different pieces of information pertinent to the crime - people, social
media check-ins, driver's licenses, crime scene reports, police
interviews, and more!

## Access the Data

This code chunk will read in **all** of the tables of data for you.
Don't modify or remove this! Take some time to look at each file type so
that

```{r}
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in
> SQL City sometime on January 15, 2018. Could you retrieve the crime
> scene report from the police department’s database and follow the
> clues to find the person responsible for the murder?!

**Step 1: Find the police report in `crime_scene_report`. Then used the
information in the report to move on to the next data source based on
the information learned.**

```{r}
#| label: relevant crime scene report

crime_scene_report |>
  filter(date == '20180115'& #on January 15, 2018
           type == "murder" & #a murder
           city == "SQL City")|>#in SQL City
  pull(description)#pull to see what the report say for valuable information

```

**Next Steps: Follow the evidence to the person responsible for the
murder, building a report as you go.** There are accomplices, some
knowingly and some unknowingly, but there is only one mastermind.

Solve the murder mystery, showing **all of your work in this document**.
Your document and code must be well organized, easy to follow, and
reproducible.

-   Use headers and written descriptions to indicate what you are doing.
-   Use `dplyr` verbs and `join` functions rather than just looking
    through the tables manually. Functions from `stringr` and
    `lubridate` will also be useful.
-   Use good code formatting practices.
-   Comment your code.
-   Cite any external sources you use to solve the mystery. \>cheat
    sheets

### Security Footage Witness Information
```{r}
#| label: information on witnesses 

#the last house on \"Northwestern Dr\

suspect_list <- bind_rows(
  person |>
    filter(address_street_name == "Northwestern Dr")|>#all the rows showing the Northwestern Dr street
    arrange(desc(address_number))|>#go from largest to smallest adress number for the "last house on the street"
    slice_head()|>
    mutate(description = "witness"),
  person|>
    filter(str_extract(name, "^[^ ]+") == "Annabel", 
           address_street_name == "Franklin Ave")|>
    mutate(description = "witness")
)

suspect_list
```

I found the syntax for [extracting after the fist
name](https://stackoverflow.com/questions/26489276/separate-name-into-firstname-and-lastname-columns-of-data-frame)

### What did Witnesses say during the Police Interviews

```{r}
#| label: police interviews
interview|>
  filter(person_id %in% c(14887, 16371))|>
  pull(transcript)
```
### Who fits Police interview Descriptions

```{r}
# Morty's(14887) suspect description

#get fit members matching witness descriptions of who ran out 
filter(get_fit_now_member,str_starts(id, "48Z") & membership_status == "gold")

#who has a license with H42W
filter(drivers_license,str_detect(plate_number, "H42W"))

#whose license matches the owners of the car that was seen 
filter(person, license_id %in% c(183779,423327,664760))

# person described having run out was Jeremy Bowers added to suspect_list
suspect_list <- bind_rows(
  suspect_list,
  person |> filter(id == 67318)|>
  mutate(description = "witness ID"))


# was Morty (14887) at the gym Facebook events on '20180115'aka January 15, 2018, and gym check in

#where were witnesses on 
facebook_event_checkin|>
  filter(person_id %in% c(14887, 16371))

#both individuals who ran out 
get_fit_now_check_in|>
  filter(membership_id %in% c('48Z7A', '48Z55') & 
           check_in_date =='20180109')


```


::: callout-note
## And the final suspect is...

*put the name of the person responsible for the murder here.*
:::
